Thermodynamic fragments in magic state theories

### 2020-10-08 magic_paper.tex
refining

\subsubsection{Binomial distributions and error bounds}
In this section, for any non-negative reals $a,b$ with $b\geq a$, we will denote the set of integers $n$ such that $a\leq n \leq b$ as $[a,b]$ for convenience.

@@

The experiment is described by binomial statistics and we define the ``left'' and ``right'' tails of the cumulative binomial distribution,
\begin{align}
	\Phi_\ell(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j}, \label{eq:phi}\\
	\Phi_r(m; n, p) &\coloneqq \sum\limits_{j=m}^n \binom{n}{j} p^j (1-p)^{n-j}, \label{eq:phir}
\end{align}
where $m\in [0,n]$.

@@

	Bound 3 follows from bound 1 and~\cref{eq:reversephir}. Bound 4 follows from bound x2 and~\cref{eq:reversephir}.

@@

We now define the alternating counterpart of the left tail distribution,
\begin{equation}\label{eq:phialt}
	\Phi_\ell^{\rm{alt}}(m; n, p) \coloneqq \sum\limits_{j=0}^m \binom{n}{j} (-p)^j (1-p)^{n-j},
\end{equation}
where $m\in [0,n]$.
$\Phi_\ell^{\rm{alt}}$ is \emph{not} the cumulative distribution of any random variable.
It is upper bounded by $\Phi_\ell$,
\begin{equation}\label{eq:alt_bound}
	\abs{\Phi_\ell^{\rm{alt}}(m; n, p)} \leq \Phi(m; n, p),\ m\in [0,n]
\end{equation}
with equality iff $m=0$.

>>>>

\subsubsection{Binomial distributions and error bounds}
\textbf{Notational convention:} For any non-negative reals $a,b$ with $b\geq a$, we will denote the set of integers $n$ such that $a\leq n \leq b$ as $[a,b]$ for convenience.\\

@@

The experiment is described by binomial statistics and we define the left tail of the cumulative binomial distribution,
\begin{equation}\label{eq:phil}
	\Phi_\ell(m; n, p) \coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j},
\end{equation}
where $m\in [0,n]$.

The symmetry between the left tail of a $p$-coin distribution and the right tail of a $(1-p)$-coin distribution dictates that
\begin{align}\label{eq:phi_reverse}
	\Phi_\ell(m; n, p) + \Phi_\ell(n-m-1; n, 1-p) &= 1,\ m\in [0,n-1],
\end{align}

Bounds on the tails of a cumulative binomial distribution can be expressed in terms of the classical relative entropy between a $p$-coin and a $q$-coin,
\begin{equation}\label{eq:ent_rel}
	\ent{p}{q} \coloneqq p \log{\frac{p}{q}} + (1-p) \log{\frac{1-p}{1-q}}.
\end{equation}
This entropy satisfies the identity
\begin{equation}
	\ent{p}{q} = \ent{1-p}{1-q}.
\end{equation}

We can now derive entropic bounds on binomial cumulative distributions~\cite{cit:ash}.
\begin{lemma}\label{lem:phil_bounds}
	Given fixed $n>0$ and $p$, $\Phi_\ell$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi_\ell(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ents},\ m\in [1,n-1] \\
		&\text{2. } \Phi_\ell(m; n, p) \leq 2^{-n\ents},\ m\in [0,np] \\
		&\text{3. } \Phi_\ell(m; n, p) \leq 1 - \left[ 8(m+1)\left(1-\frac{m+1}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m+1}{n}}{p}}, \\
		&\hspace{14pt} m\in [0,n-2]
		\end{split}
		\\
		&\text{4. } \Phi_\ell(m; n, p) \geq 1 - 2^{-n\ent{\frac{m+1}{n}}{p}},\ m\in [np+1,n-2]
	\end{align*}
\end{lemma}

We now define the partial sums
\begin{align}	
	\Phi_+(m; n, p) &\coloneqq \sum\limits_{j=0}^{m/2} \binom{n}{2j} p^{2j} (1-p)^{n-2j}, \nonumber\\ 
	&m\in[0,n]\ \text{and m even}, \label{eq:fp} \\
	\Phi_-(m; n, p) &\coloneqq \sum\limits_{j=1}^{(m-1)/2} \binom{n}{2j+1} p^{2j+1} (1-p)^{n-(2j+1)}, \nonumber\\ 
	&m\in[0,n]\ \text{and m odd}. \label{eq:fn2}
\end{align}
Index $m$ only takes even (odd) values when labelling $\Phi_+$ ($\Phi_-$) and therefore we can use $\Phi_{\pm}$ to compactify statements that hold for both sums.
In particular, we derive ``loose'' and ``strict'' bounds on $\Phi_{\pm}$.
\begin{lemma}\label{lem:bounds_loose}
	Given fixed $n>0$ and $p$, $\Phi$ satisfies the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi_{\pm}(m; n, p) \geq \left[ 8m\left(1-\frac{m}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ents},\ m\in [1,n-1] \\
		&\text{2. } \Phi_{\pm}(m; n, p) \leq 2^{-n\ents},\ m\in [0,np] \\
		&\text{3. } \Phi_{\pm}(m; n, p) \leq 1 - \left[ 8(m+1)\left(1-\frac{m+1}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{m+1}{n}}{p}}, \\
		&\hspace{14pt} m\in [0,n-2]
		\end{split}
	\end{align*}
\end{lemma}
\begin{proof}
	!!!
\end{proof}
\begin{lemma}\label{lem:bounds_strict}
	Given fixed $n>0$ and $p$, $\Phi_+, \Phi_-$ satisfy the following bounds:
	\begin{align*}
		\begin{split}
		&\text{1. } \Phi_+(m; n, p) \geq \sum\limits_{j=0}^{m/2}\left[ 8(2j)\left(1-\frac{2j}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{2j}{n}}{p}}, \\
		&\hspace{14pt} \text{for all even } m\in [2,n] \\
		&\text{2. } \Phi_+(m; n, p) \leq \sum\limits_{j=0}^{m/2}\left[ 2\pi(2j)\left(1-\frac{2j}{n}\right) \right]^{-\frac{1}{2}} 2^{-n\ent{\frac{2j}{n}}{p}}, \\
		&\hspace{14pt} \text{for all even } m\in [2,n] \\
		&\text{3. } \Phi_-(m; n, p) \geq \sum\limits_{j=1}^{(m-1)/2}\left[ 8(2j+1)\left(1-\frac{2j+1}{n}\right) \right]^{-\frac{1}{2}} \times \\
		&\hspace{14pt} \times 2^{-n\ent{\frac{2j+1}{n}}{p}},\ \text{for all odd } m\in [1,n] \\
		&\text{4. } \Phi_-(m; n, p) \leq \sum\limits_{j=1}^{(m-1)/2}\left[ 2\pi(2j+1)\left(1-\frac{2j+1}{n}\right) \right]^{-\frac{1}{2}} \times \\
		&\hspace{14pt} \times 2^{-n\ent{\frac{2j+1}{n}}{p}},\ \text{for all odd } m\in [1,n]
		\end{split}
	\end{align*}
\end{lemma}
\begin{proof}
	!!!
\end{proof}

The experiment is described by binomial statistics and we define the left and ``right'' tails of the cumulative binomial distribution,
\begin{align}
	\Phi_\ell(m; n, p) &\coloneqq \sum\limits_{j=0}^m \binom{n}{j} p^j (1-p)^{n-j}, \label{eq:phi}\\
	\Phi_r(m; n, p) &\coloneqq \sum\limits_{j=m}^n \binom{n}{j} p^j (1-p)^{n-j}, \label{eq:phir}
\end{align}
where $m\in [0,n]$.

@@

	Bound 3 follows from bound 1 and~\cref{eq:reversephir}. Bound 4 follows from bound 2 and~\cref{eq:reversephir}.

@@

We now define the alternating counterpart of the left tail distribution,
\begin{equation}\label{eq:phialt}
	\Phi_\ell^{\rm{alt}}(m; n, p) \coloneqq \sum\limits_{j=0}^m \binom{n}{j} (-p)^j (1-p)^{n-j},
\end{equation}
where $m\in [0,n]$.
$\Phi_\ell^{\rm{alt}}$ is \emph{not} the cumulative distribution of any random variable.
It is upper bounded by $\Phi_\ell$,
\begin{equation}\label{eq:phialt_bound}
	\abs{\Phi_\ell^{\rm{alt}}(m; n, p)} \leq \Phi_\ell(m; n, p),\ m\in [0,n]
\end{equation}
with equality iff $m=0$.
